Understanding the nanoscale structure of hexagonal phase lyotropic liquid crystal membranes

### 2017-12-19 Ben_Manuscripts/structure_paper/Draft.tex
updated draft with comments.

\documentclass{article}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{subcaption}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath} % or simply amstext
\usepackage{siunitx}
\usepackage{booktabs}
\usepackage[export]{adjustbox}
\newcommand{\angstrom}{\textup{\AA}}
\usepackage{cleveref}
\usepackage{booktabs}
\usepackage{gensymb}
\usepackage{float}
\usepackage{setspace}
\doublespacing

@@

@@

  %BJC : I don't know that this paragraph fits.  
  %MRS3: not really. Maybe take a look at Osuji's recent review (2016, I think) that addresses applications of nanostructured membranes and uses.  Could rewrite the introduction to be more about the general possibilities of nanostructured membranes, not just desalination.  It should be an introduction to what you will talk about in the paper, not to everything related.
 % Improved RO membranes can make desalination more sustainable. Only 0.5 \% of
 % the world's water is fresh. Desalination is an important source of potable
 % water, necessary in order to keep up with demand generated by a rapidly growing
 % global population. Although RO is currently the cheapest form of desalination,
 % the cost must be further decreased in order to meet demand. A typical seawater
 % desalination process requires feed pressures in the range of 800 to 1000 psi
 % \cite{fritzmann_state---art_2007}. By designing better membranes for
 % desalination, we can achieve higher water fluxes and reduce feed pressure
 % requirements which will decrease the cost of fresh water production.
 % \cite{humplik_nanostructured_2011}.

@@

  The permeability-selectivity tradeoff has the potential to be overcome by
  designing membranes at the molecular level. Next-geneartion nanostructured
  membranes with high selectivity permitted by a precisely controlled pore size
  and high permeability allowed by its porous architecture, have the potential to
  replace traditional RO and NF membrane technologies.

@@

  Self assembling lyotropic liquid crystals (LLCs) are a suitable candidate for
  aqueous separation applications. LLCs share the characteristic ability of
  nanostructured membrane materials to create highly ordered structures with the
  added benefits of low cost and synthetic techniques feasible for large scale
  production.  \cite{feng_scalable_2014} LLC systems created by the monomer 
  Na-GA3C11 (Fig.~\ref{fig:monomer}) have been extensively studied experimentally.
  \cite{smith_ordered_1997,zhou_supported_2005,resel_h2-phase_2000,feng_scalable_2014,feng_thin_2016}
  Neat liquid crystal monomer forms the thermotropic, Col\textsubscript{h}
  phase (Fig.\ref{fig:assembly}). The presence of ca. 10 wt \% water results in the
  H\textsubscript{II} phase. In both cases, monomers assemble into mesophases
  made of hexagonally packed, uniform size, cylinders with hydrophilic head
  groups oriented inward towards the cylinder center.  The hydrophilic region can
  act as a pore for aqueous separations.  One can envision tailoring the pore
  region for specific separations by changing the monomer chemistry.

  Research into LLC membranes has been revived in recent years. During early
  stages of exploration, mesophases formed by Na-GA3C11 could not be
  macroscopically aligned, resulting in low flux membranes, and no clear route
  towards scalable and economical filtration. In 2014, Feng et al. showed that
  the mesophases could be aligned using a magnetic field with subsequent
  crosslinking to lock the structure in place \cite{feng_scalable_2014}. In 2016,
  Feng et al. showed that the same result could be obtained using a technique
  termed soft confinement \cite{feng_thin_2016}.

@@

  Our current understanding of LLC systems is not rich enough to be able to
  precisely design membranes for specific separations. Over the past 20 years,
  H\textsubscript{II} phase LLC membrane studies have been limited primarily to
  Na-GA3C11 with some characterization done after minor structural modifications.
  Resel et al. varied the length of the monomer tails and the counterion used and
  observed its affect on pore spacing \cite{resel_structural_2000}.  Their study
  offers no insight into dynamics within the pore. In a later study of rejection
  performance, it was shown that membranes formed by Na-GA3C11 cannot perform
  separations of solutes less than 1.2 nm in diameter because the pores are too
  large \cite{zhou_supported_2005}.  We do not yet understand how to controllably
  reduce the effective pore size or how to tune the chemical environment in the
  nanopores for effective water desalination and small organic molecule
  separations. The only source of predictive modeling for LLC systems have been
  macroscopic models which likely do not adequately describe transport at these
  length scales \cite{hatakeyama_water_2011}. It will be challenging to
  efficiently narrow down the large design space in a laboratory setting without
  a robust model.

@@

  A molecular level understanding of LLC membrane structure, enabled by
  molecular dynamics simulations, can provide guidelines to reduce the large
  chemical space available to design monomers for creation of separation-specific
  membranes. A good molecular model should incorporate a detailed picture of the
  nanoscopic pore structure which will be crucial to understanding the role of
  monomer structure in solute transport and membrane design. Models resulting
  from molecular dynamics simulations will provide the required level of detail
  (Fig. \ref{fig:detail}).  We can directly observe solute transport and suggest
  governing mechanisms. We can observe how the choice of head group may influence
  pore size for size exclusion driven separations. We can interchange counterions
  which may influence both the pore size and the strength of the Donnan potential
  which affects the degree to which the membrane can exclude charged species.

@@

  %BJC: I think the first sentence of the following paragraph is
  %necessary to put somewhere, but maybe not in this spot exactly
  In order to appropriately model transport, we must first gain a thorough
  understanding of the nanoscopic structure of LLC membranes. Our approach to
  constructing a general model will follow the development of a model of the
  assembly formed by Na-GA3C11 since it has sufficient experimental
  characterization. We have also narrowed our scope to the development of a model
  of the Col\textsubscript{h} phase membrane. Compared to the H\textsubscript{II}
  phase, the Col\textsubscript{h} phase is a simpler starting point, due to the
  absence of water, and has detailed experimental wide-angle X-ray scattering
  (WAXS) patterns useful for reconstructing structural data.

  Despite having structural data, there is still information which experiment
  cannot definitively answer. There are several key questions that we will
  investigate.

  Monomers in the Col\textsubscript{h} system are theorized to be partitioned
  into stacked layers which form columnar pores. We want to know

@@

  A simple molecular simulation study of a similar molecule suggested that
  there are 4 monomers in each layer. Their estimation is based on a simulated
  system containing only 16 total monomers which likely does not sufficiently
  model the chemical environment present in the real system
  \cite{zhu_methacrylated_2006}.  A separate calculation based on the volume of
  the liquid crystal monomers proposes that there are seven monomers in each
  layer \cite{resel_structural_2000}.  A molecular model orders of magnitude
  larger than any other reported atomistic liquid crystal membrane simulations
  has the best chance of directly answering this question.  We will directly
  change the layer composition and note its effect on membrane structure.

@@

  The $\pi$-$\pi$ stacking interactions may be a driving force of self assembly
  in this system \cite{gazit_possible_2002}. Gas phase ab initio studies of
  benzene dimers have shown a clear energetic advantage for parallel displaced
  and T-shaped $\pi$-$\pi$ stacking conformations versus a sandwiched
  conformation ~\cite{sinnokrot_estimates_2002}. Substituted benzene rings
  exhibit an even stronger $\pi$-$\pi$ stacking attraction which favors the
  parallel displaced configuration in all cases except where the substitutions
  are extremely electron withdrawing.
  \cite{waller_hybrid_2006,ringer_effect_2006}. In this study, we compare
  simulated X-ray diffraction patterns to experiment in order to deduce which
  stacking configurations is most likely.

@@

  While the Col\textsubscript{h} phase is described as dry, it has been
  suggested by experimentalists, in unpublished communications, that it is likely
  that small amounts of ambient water are leached into neat monomer.
  Experimentally, achieving a hexagonal phase with a completely dry system has
  proven difficult. If neat monomer is allowed to sit in ambient conditions, its
  color turns from transparent to slightly opaque and a hexagonal phase forms.
  Although we will not explore whether water is necessary for self assembly, we
  hypothesize that the hydrogen bonding network formed by the water may play a
  role in structuring the pores and holding together the hexagonal phase. We can
  use simulated X-ray diffraction patterns to see if there is any meaningful
  structural difference between a ``dry'' and ``wet'' system.

@@

  %BJC: TODO: update colorbar
  \begin{figure}
        \centering
        \begin{subfigure}[t]{0.43\linewidth}
                \centering
                \includegraphics[width=\linewidth]{SAXS.png}
                \caption{}\label{fig:SAXS}
        \end{subfigure}
        \begin{subfigure}[t]{0.47\linewidth}
                \centering
                \includegraphics[width=\linewidth]{WAXS_annotated.png} 
                \caption{}\label{fig:WAXS}
        \end{subfigure}
%MRS3: would be good to have all feature labels agree: either can edit to put the names on, or have a key relating numbers to the names you use.
%BJC2: I will annotate it with the names R-pi, R-alkanes etc. Is that what you mean?
	\caption{(a) The repeat spacing in the 1D small angle X-ray scattering pattern
		is characteristic of hexagonal packing. The leading peak represents the
		distance between the d\textsubscript{100} planes. Using this distance, we know
		that the distance between pore centers is 4.12 nm. (b) 2D wide angle X-ray
		scattering gives details about repeating features less than 1 nanometer
		apart. Experimentalists have justified each of the 5 major reflections
		present as follows: (I) Aromatic head group $\pi-\pi$ stack 3.7 \AA~apart.
		(II) Monomers arrange vertically in a 2\textsubscript{1} helix. (III) Alkane
 		chain tails pack 4.5 \AA~apart. (IV) Monomer tails tilt with respect to the 
 		membrane plane. (V) As derived from SAXS, pores pack hexagonally and are 
		spaced 4.12 nm apart}
	\label{fig:SWAXS}
 \end{figure}

@@

  Liquid crystal monomers were parameterized using the Generalized AMBER
  Forcefield \cite{wang_development_2004} with the Antechamber package
  \cite{wang_automatic_2006} provided with AmberTools16
  \cite{case_ambertools16_2016}. Atomic charges were assigned using the am1bccsym
  method of molcharge shipped with QUACPAC from Openeye Scientific Software. All
  molecular dynamics simulations were run using Gromacs 2016.
  \cite{bekker_gromacs:_1993,berendsen_gromacs:_1995,van_der_spoel_gromacs:_2005,hess_gromacs_2008}

  An ensemble of characteristic, low-energy vacuum monomer configurations
  were constructed by applying a simulated annealing process to a parameterized
  monomer. Monomers were cooled from 1000K to 50K over 10 nanoseconds.  A low
%MRS3: need to describe this in more detail, but can be in the supplementary. 
%BJC2: would it be best to include the parameterization script
  energy configuration was randomly pulled from the trajectory and charges were
  reassigned using molcharge.
%MRS3: need to say what molcharge is.
%BJC2: I say in paragraph above : "Atomic charges were assigned using the am1bccsym method of 
%molcharge shipped with QUACPAC from ...". Is that not enough?
  Using the new charges, the monomer system was
  annealed again and a random monomer configuration was pulled from the
  trajectory to be used for full system construction (Figure~\ref{fig:python}a).
  See supplemental information for further detail.

@@

  The timescale for self assembly of monomers into the hexagonal phase is
  unknown and likely outside of a reasonable length for an atomistic simulation,
  calling for a more efficient way to build the system.  Previous work has shown
  a coarse grain model self assemble into the H\textsubscript{II} phase
  configuration in $\approx$ 1000 ns \cite{mondal_self-assembly_2013}.  We
  attempted atomistic self-assembly by packing monomers into a box using Packmol
  \cite{martinez_packmol:_2009}.  Simulations of greater than 100 ns show no
  indicators of progress towards an ordered system To bypass the slow
  self-assembly process, python scripts are used to assemble monomers into a
  structure close to one of a number of hypothesized equilibrium configurations
  (Figure~\ref{fig:python}).

@@

  When constructing an initial configuration, there are a number of variables
  which require careful consideration while placing monomers. The equilibrium
  configuration is sensitive to some while insensitive to others. The starting
  pore radius, defined as the distance of a chosen head group carbon from the
  pore's central axis, does not influence the equilibrium structure when a
  reasonable value is chosen (See Supplemental). The pore radius is chosen to be
  0.6 nm in our initial configurations because the pore size is estimated to be
  $\approx$ 1.2 nm. The initial distance between pores also has little effect on
  the the equilibrated structure. However, one should not start them too close or
  there will be high energy repulsions during early equilibration. We chose an
  initial pore spacing of 4.5 nm, $\approx$ 10 \% larger than the experimental
  value of 4.12 nm.  The distance between layers, the rotation of the layers with
  respect to adjacent layers, and the number of monomers per layer do influence
  the equilibrium structure and require further justification for their choices.
  We rely on experimental data to inform them.

  We chose the layer spacing for the initial configuration based on R-$\pi$ and
  then allowed the system to readjust during equilibration. Each monomer was rotated
  so the plane of the aromatic head groups would be coplanar with the xy plane.
  We explore two different initial layer spacings. The first is exactly equal to
  R-$\pi$ with layers placed so aromatic rings are stacked 3.7 \angstrom~ apart
  in the z-direction. A second system is explored with an initial layer spacing
  of 5 \angstrom. A third system with an initial layer spacing of 10 \angstrom was
  briefly explored. When layers are spaced out too far, they will collapse
  on each other while simultaneously slipping in between layers of adjacent pores
  which leads to an artificially thick membrane with pores spaced closely
  together. 
  Further details of simulations with large layer spacing are in the supplemental information.
%There will be no further discussion of this system but the interested
%  reader can learn more about it in the supplemental information.

  The relative interlayer orientation was chosen based on clues from
  diffraction data as well as the various known stacking modes of benzene and
  substituted benzene rings: sandwiched, parallel-displaced and T-shaped
  ~\cite{sinnokrot_estimates_2002} (\Cref{fig:sandwiched,fig:pd,fig:tshaped}).
  The T-shaped configuration was ruled out because its $\approx$ 5 \angstrom~
  equilibrium stacking distance ~\cite{sinnokrot_estimates_2002} is inconsistent
  with R-$\pi$. It is also infeasible for the monomers to orient in the T-shaped
  conformation because of the bulky tail groups. The system's preference towards
  the sandwiched vs. parallel displaced stacking modes will be explored in some
  detail.  Both have reported stacking distances near the R-$\pi$ value of 3.7
  \angstrom. Headgroups in our sandwiched initial configuration are stacked
  directly on top of each other while stacked headgroups in the parallel
  displaced initial configuration are offset by $180/nmon$ degrees where $nmon$
  equals the number of monomers per layer.

  As outlined in (\ref{point:monomernum}) the number of monomers in each layer is
  unknown. We tested configurations constructed with a varied number of
  monomers per layer. Systems were built in the offset and parallel displaced
  configurations with 4, 5, 6, 7 and 8 monomers per layer.

@@

  Using equilibrated structures, we carry out various calculations to
  characterize the system. We define the point at which a system is
  equilibrated based on when the distance between pores stops changing.
  We determined when the distances stopped changing by applying the statistical
  test, \texttt{pymbar.timeseries.detectEquilibration}, to the time series. 
  Simulations of 400 ns give at least 50 ns of equilibrated simulation trajectory.
  %BJC2: Do I cite the github repository for pymbar?

  To calculate the equilibrated pore spacing, we measured the distance between
  pore centers. Pore centers are located by averaging the coordinates of sodium
  ions in their respective pores. Pore spacing statistics were generated 
  using the bootstrapping technique (See Supplemental Information).

@@

  % BJC: To be replaced by Joe's description
  Simulated X-ray diffraction patterns are generated based on atomic
  coordinates in order to make a direct experimental comparison. All atomic
  coordinates were simulated as Gaussian spheres of electron density
  corresponding to each atom's atomic number. A three dimensional Fourier
  transform (FT) of the array of electron density results in a three dimensional
  structure factor which represents the unit cell in reciprocal space. We matched
  experimental 2D WAXS patterns by adjusting the initial spacing between layers
  and the orientation of the head groups with respect to adjacent layers.

  The colorbars on all diffraction patterns are normalized relative to
  R-alkanes. We believe that the alkane-alkane density, averaged over all angles,
  is the feature most likely to be replicated between experiment and simulation.
  Other features are dependent on system ordering which is likely to have some
  dependence on initial configuration. We calculated the average intensity within
  R-alkanes of the experimental pattern, $I_{avg}$.  We exclude intensities
  within $\pm$ 30\degree~of the $q_r=0$ axis, since the simulated patterns differ
  from experiment in those regions. In contrast to the experimental WAXS pattern,
  R-$\pi$ appearing in simulated diffraction patterns intersects with R-alkanes
  (See Fig. \ref{fig:XRDsim}). We multiplied $I_{avg}$ by a scaling factor of
  2.5. Intensities that appear in the experimental pattern $\geq$ 2.5$*I_{avg}$
  receive colorbar values of 1. We apply the same scaling method to the simulated
  patterns. We carefully chose a scaling factor of 2.5 in order to (1) visibly
  display all features in the experimental pattern and (2) to allow us to compare
  the relative intensities of features between simulated and experimental
  diffraction patterns.

>>>>

\documentclass{article}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{subcaption}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath} % or simply amstext
\usepackage{siunitx}
\usepackage{booktabs}
\usepackage[export]{adjustbox}
\newcommand{\angstrom}{\textup{\AA}}
\usepackage{cleveref}
\usepackage{booktabs}
\usepackage{gensymb}
\usepackage{float}
\usepackage{setspace}
\usepackage{color}
\doublespacing

@@

  %MRS4: are all commercial RO membranes just like this? This seems
  %awfully specific.  
  %MRS4: these two paragraphs are perhaps not
  %ideal.  They describe the process of making the membranes, but that
  %is not what we need as the introduction to THIS paper.  Instead, you
  %need to focus on the strength/issues with RO and NF that motivate this research.

@@

  %BJC : I don't know that this paragraph fits.  
  %MRS3: not really. Maybe take a look at Osuji's recent review (2016, I think) that addresses applications of nanostructured membranes and uses.  Could rewrite the introduction to be more about the general possibilities of nanostructured membranes, not just desalination.  It should be an introduction to what you will talk about in the paper, not to everything related.
 % Improved RO membranes can make desalination more sustainable. Only 0.5\% of
 % the world's water is fresh. Desalination is an important source of potable
 % water, necessary in order to keep up with demand generated by a rapidly growing
 % global population. Although RO is currently the cheapest form of desalination,
 % the cost must be further decreased in order to meet demand. A typical seawater
 % desalination process requires feed pressures in the range of 800 to 1000 psi
 % \cite{fritzmann_state---art_2007}. By designing better membranes for
 % desalination, we can achieve higher water fluxes and reduce feed pressure
 % requirements which will decrease the cost of fresh water production.
 % \cite{humplik_nanostructured_2011}.

@@

  The permeability-selectivity tradeoff has the potential to be overcome by
  designing membranes at the molecular level. Next-generation nanostructured
  membranes with high selectivity permitted by a precisely controlled pore size
  and high permeability allowed by its porous architecture, have the potential to
  replace traditional RO and NF membrane technologies.

@@

  Self assembling lyotropic liquid crystals (LLCs) are a suitable candidate for
  aqueous separation applications. LLCs share the characteristic ability of
  nanostructured membrane materials to create highly ordered structures with the
  added benefits of low cost and synthetic techniques feasible for large scale
  production.  \cite{feng_scalable_2014} LLC systems created by the monomer 
  Na-GA3C11 (Fig.~\ref{fig:monomer}) have been extensively studied experimentally.
  \cite{smith_ordered_1997,zhou_supported_2005,resel_h2-phase_2000,feng_scalable_2014,feng_thin_2016}
  Neat liquid crystal monomer forms the thermotropic, Col\textsubscript{h}
  phase (Fig.\ref{fig:assembly}). The presence of ca. 10\% wt water results in the
  H\textsubscript{II} phase. In both cases, monomers assemble into mesophases
  made of hexagonally packed, uniform size, cylinders with hydrophilic head
  groups oriented inward towards the cylinder center.  The hydrophilic region can
  act as a pore for aqueous separations. 
  %MRS4: aren't' there some citations where Gin and noble at least try to separate things (Even at low efficiency) to determine the size? (zhou_supported_2005) that should be cited here on ``act as a pore'' to show these are at least existing (if not usable
 One can envision tailoring the pore
  region for specific separations by changing the monomer chemistry.
%MRS4: can cite the Resel paper paper when changed tails and got different pore-to-pore distances as an example of changing monomer chemistry.

  Research into LLC membranes has been revived in recent years. During early
  stages of exploration, mesophases formed by Na-GA3C11 could not be
  macroscopically aligned, resulting in low flux membranes, and no clear route
  towards scalable and economical filtration. In 2014, Feng et al.~showed that
  the mesophases could be aligned using a magnetic field with subsequent
  crosslinking to lock the structure in place \cite{feng_scalable_2014}. In 2016,
  Feng et al.~showed that the same result could be obtained using a technique
  termed soft confinement \cite{feng_thin_2016}.

@@

  Our current understanding of LLC systems is not rich enough to be able to
  precisely design membranes for specific separations. Over the past 20 years,
  H\textsubscript{II} phase LLC membrane studies have been limited primarily to
  Na-GA3C11 with some characterization done after minor structural modifications.
  Resel et al.~varied the length of the monomer tails and the counterion used and
  observed its affect on pore spacing \cite{resel_structural_2000}.

%MRS4: is this sentence needed here? was there an expectation it would? 
%Their study offers no insight into dynamics within the pore. 
In a later study of rejection
  performance, it was shown that membranes formed by Na-GA3C11 cannot perform
  separations of solutes less than 1.2 nm in diameter because the pores are too
  large \cite{zhou_supported_2005}.  We do not yet understand how to controllably
  reduce the effective pore size or how to tune the chemical environment in the
  nanopores for effective water desalination and small organic molecule
  separations. The only source of predictive modeling for LLC systems have been
  macroscopic models which likely do not adequately describe transport at these
  length scales \cite{hatakeyama_water_2011}. It will be challenging to
  efficiently narrow down the large design space in a laboratory setting without
  a robust model.

@@

  A molecular level understanding of LLC membrane structure, enabled by
  molecular dynamics simulations, can provide guidelines to reduce the large
  chemical space available to design monomers for creation of separation-specific
  membranes. A good molecular model should incorporate a detailed picture of the
  nanoscopic pore structure which will be crucial to understanding the role of
  monomer structure in solute transport and membrane design. Models resulting
  from molecular dynamics simulations will provide the required level of detail
  (Fig. \ref{fig:detail}),
%MRS4: added as caveat.
  assuming the force fields are sufficiently accurate.
%MRS4: also added.
  with such an atomistic model,
  We can directly observe solute transport and suggest
  governing mechanisms. We can observe how the choice of head group may influence
  pore size for size exclusion driven separations. We can interchange counterions
  which may influence both the pore size and the strength of the Donnan potential
  which affects the degree to which the membrane can exclude charged species.

@@

  %BJC: I think the first sentence of the following paragraph is
  %necessary to put somewhere, but maybe not in this spot exactly
  In order to appropriately model transport, we must first gain a thorough
  understanding of the nanoscopic structure of LLC membranes. Our approach to
  constructing a general model will follow the development of a model of the
  assembly formed by Na-GA3C11 since it has sufficient experimental
  characterization. We have also narrowed our scope to the development of a model
  of the Col\textsubscript{h} phase membrane. Compared to the H\textsubscript{II}
  phase, the Col\textsubscript{h} phase is a simpler starting point, due to the
  absence of water, and has detailed experimental wide-angle X-ray scattering
  (WAXS) patterns useful for reconstructing structural data.  
  %MRS4: you should mention that the two phases appear to have similar pore-to-pore distances, 
  %suggesting they are not very different.

  %MRS4: weird to have a one sentence paragraph.  Maybe combine with next after rephrasing?
  Despite having structural data, there is still information which experiment
  cannot definitively answer. There are several key questions that we will
  investigate.

  Monomers in the Col\textsubscript{h} system are theorized to be partitioned
  into stacked layers which form columnar pores. We %MRS4:want 
  would like to know

@@

  A simple molecular simulation study of a similar molecule suggested that
  there are 4 monomers in each layer.!\cite{zhu_methacrylated_2006}.  
  %MRS4: moving citation to end of first reference of a paper. 
%MRS4: rephrase:
However, their estimate 
is based on a simulated
  system containing only 16 total monomers which likely does not sufficiently
  model the chemical environment present in the real system

A separate calculation based on the volume of
  the liquid crystal monomers proposes that there are seven monomers in each
  layer \cite{resel_structural_2000}.  A molecular model orders of magnitude
  larger than any other reported atomistic liquid crystal membrane simulations,
%MR4: added 
as we present here,
  has the best chance of directly answering this question.  We will directly
  change the layer composition and note its effect on membrane structure.

@@

  The $\pi$-$\pi$ stacking interactions may be a driving force of self-assembly
  in this system \cite{gazit_possible_2002}. Gas phase ab initio studies of
  benzene dimers have shown a clear energetic advantage for parallel displaced
  and T-shaped $\pi$-$\pi$ stacking conformations versus a sandwiched
  conformation ~\cite{sinnokrot_estimates_2002}. Substituted benzene rings
  exhibit an even stronger $\pi$-$\pi$ stacking attraction which favors the
  parallel displaced configuration in all cases except where the substitutions
  are extremely electron withdrawing.
  \cite{waller_hybrid_2006,ringer_effect_2006}. In this study, we compare
  simulated X-ray diffraction patterns to experiment in order to deduce which
  stacking configurations is most likely.

@@

  While the Col\textsubscript{h} phase is described as dry, it has been
  suggested by experimentalists, in unpublished communications, that it is likely
  that small amounts of ambient water are leached into neat monomer.
  Experimentally, achieving a hexagonal phase with a completely dry system has
  proven difficult. If neat monomer is allowed to sit in ambient conditions, its
  color turns from transparent to slightly opaque and a hexagonal phase forms.
  Although we will not explore whether water is necessary for self-assembly, we
  hypothesize that the hydrogen bonding network formed by the water may play a
  role in structuring the pores and holding together the hexagonal phase. We can
  use simulated X-ray diffraction patterns to see if there is any meaningful
  structural difference between a ``dry'' and ``wet'' system.

@@

  %BJC: TODO: update colorbar
  \begin{figure}
        \centering
        \begin{subfigure}[t]{0.43\linewidth}
                \centering
                \includegraphics[width=\linewidth]{SAXS.png}
                \caption{}\label{fig:SAXS}
        \end{subfigure}
        \begin{subfigure}[t]{0.47\linewidth}
                \centering
                \includegraphics[width=\linewidth]{WAXS_annotated.png} 
                \caption{}\label{fig:WAXS}
        \end{subfigure}
%MRS3: would be good to have all feature labels agree: either can edit to put the names on, or have a key relating numbers to the names you use.
%BJC2: I will annotate it with the names R-pi, R-alkanes etc. Is that what you mean?
%MRS4: that works fine.  We just need a way to connect what is in the figure with the names used in the text.  Ideally there would not be two sets of labels, just one. 
	\caption{(a) The repeat spacing in the 1D small angle X-ray scattering pattern
		is characteristic of hexagonal packing. The leading peak represents the
		distance between the d\textsubscript{100} planes. Using this distance, we know
		that the distance between pore centers is 4.12 nm. (b) 2D wide angle X-ray
		scattering gives details about repeating features less than 1 nanometer
		apart. Experimentalists have justified each of the 5 major reflections
		present as follows: (I) Aromatic head group $\pi-\pi$ stack 3.7 \AA~apart.
		(II) Monomers arrange vertically in a 2\textsubscript{1} helix. (III) Alkane
 		chain tails pack 4.5 \AA~apart. (IV) Monomer tails tilt with respect to the 
 		membrane plane. (V) As derived from SAXS, pores pack hexagonally and are 
		spaced 4.12 nm apart}
	\label{fig:SWAXS}
 \end{figure}

@@

  Liquid crystal monomers were parameterized using the Generalized AMBER
  Forcefield \cite{wang_development_2004} with the Antechamber package
  \cite{wang_automatic_2006} provided with AmberTools16
  \cite{case_ambertools16_2016}. Atomic charges were assigned using the am1bccsym
  method of molcharge shipped with QUACPAC from Openeye Scientific Software. All
  molecular dynamics simulations were run using Gromacs 2016.
  \cite{bekker_gromacs:_1993,berendsen_gromacs:_1995,van_der_spoel_gromacs:_2005,hess_gromacs_2008}
%MRS4: somewhere would need to specify things like time step tau_t, tau_p, etc., so that people could in theory replicate 
the experiments.  Note: we should put scripts and starting configurations (not all that were done but the important ones) in a public repository by the time of submission.  
  An ensemble of characteristic, low-energy vacuum monomer configurations
  were constructed by applying a simulated annealing process to a parameterized
  monomer. Monomers were cooled from 1000K to 50K over 10 nanoseconds.  A low
%MRS3: need to describe this in more detail, but can be in the supplementary. 
%BJC2: would it be best to include the parameterization script
  energy configuration was randomly pulled from the trajectory and charges were
  reassigned using molcharge.
%MRS4: I would give a ``plain text'' version of the script in supplementary.
%MRS3: need to say what molcharge is.
%BJC2: I say in paragraph above : "Atomic charges were assigned using the am1bccsym method of 
%molcharge shipped with QUACPAC from ...". Is that not enough?
%MRS4: sorry, missed that.  Maybe put software command names in \texttt{software} to make it clear what they are.
  Using the new charges, the monomer system was
  annealed again and a random monomer configuration was pulled from the
  trajectory to be used for full system construction (Figure~\ref{fig:python}a).
  See supplemental information for further detail.

@@

%MRS4: check grammar additions in this paragraph.
  The timescale for self-assembly of monomers into the hexagonal phase is
  unknown and likely outside of a reasonable length for an atomistic simulation,
  calling for a more efficient way to build the system.  Previous work has shown
  a coarse-grained model self assemble into the H\textsubscript{II} phase
  configuration in $\sim$1000 ns \cite{mondal_self-assembly_2013}.  We
  attempted atomistic self-assembly by packing monomers into a box using Packmol
  \cite{martinez_packmol:_2009}.  Simulations of greater than 100 ns show no
  indicators of progress towards an ordered system. To bypass the slow
  self-assembly process, Python scripts are used to assemble monomers into a
  structure close to one of a number of hypothesized equilibrium configurations
  (Figure~\ref{fig:python}).

@@

  When constructing an initial configuration, there are a number of variables
  which require careful consideration while placing monomers. The equilibrium
  configuration is sensitive to some while insensitive to others. The starting
  pore radius, defined as the distance of a chosen head group carbon from the
  pore's central axis, does not influence the equilibrium structure when a
  reasonable value is chosen (See Supplemental Information). The pore radius is chosen to be
  0.6 nm in our initial configurations because the pore size is estimated to be
  $\approx$ 1.2 nm.
%MRS4: provide citation here for 1.2 number.
 The initial distance between pores also has little effect on
  the the equilibrated structure. However, one should not start them too close or
  there will be high energy repulsions during early equilibration. We chose an
  initial pore spacing of 4.5 nm, $\sim$10\% larger than the experimental
  value of 4.12 nm.  The distance between layers, the rotation of the layers with
  respect to adjacent layers, and the number of monomers per layer do influence
  the equilibrium structure and require further justification for their choices.
  We rely on experimental data to inform them.

  We chose the layer spacing for the initial configuration based on R-$\pi$ and
  then allowed the system to readjust during equilibration. Each monomer was rotated
  so the plane of the aromatic head groups would be coplanar with the xy plane.
  We explore two different initial layer spacings. The first is exactly equal to
  R-$\pi$ with layers placed so aromatic rings are stacked 3.7 \angstrom~apart
  in the z-direction. A second system is explored with an initial layer spacing
  of 5 \angstrom. A third system with an initial layer spacing of 10 \angstrom~was
  %MRS4: this is probably OK here, but you do want to move away from describing things historically (the order in which the ideas were had and tested) and the logical ordering to communicate the story.  You want to do the second whenever possible. Think about this going over the paper and try to shift from historical presentation of results to story-based ordering.
  briefly explored. When layers are spaced out too far, they will collapse
  on each other while simultaneously slipping in between layers of adjacent pores,
  which leads to an artificially thick membrane with pores spaced closely
  together. 
  Further details of simulations with large layer spacing are in the supplemental information.
%There will be no further discussion of this system but the interested
%  reader can learn more about it in the supplemental information.

  The relative interlayer orientation was chosen based on clues from
  diffraction data as well as the various known stacking modes of benzene and
  substituted benzene rings: sandwiched, parallel-displaced and T-shaped
  ~\cite{sinnokrot_estimates_2002} (\Cref{fig:sandwiched,fig:pd,fig:tshaped}).
  The T-shaped configuration was ruled out because its $\sim$5 \angstrom~
  equilibrium stacking distance ~\cite{sinnokrot_estimates_2002} is inconsistent
  with R-$\pi$. It is also infeasible for the monomers to orient in the T-shaped
  conformation because of the bulky tail groups. The system's preference towards
  the sandwiched vs. parallel displaced stacking modes will be explored in some
  detail.  Both have reported stacking distances near the R-$\pi$ value of 3.7
  \angstrom. Headgroups in our sandwiched initial configuration are stacked
  directly on top of each other while stacked headgroups in the parallel
  displaced initial configuration are offset by $180/nmon$ degrees where $nmon$
  equals the number of monomers per layer.

%MRS4: describe what the points are, people don't want to go back and look.
  As outlined in (\ref{point:monomernum}) the number of monomers in each layer is
  unknown. We tested configurations constructed with a varied number of
  monomers per layer. Systems were built in the offset and parallel displaced
  configurations with 4, 5, 6, 7 and 8 monomers per layer.

@@

  Using equilibrated structures, we carry out various calculations to
  characterize the system. We define the point at which a system is
  equilibrated based on when the distance between pores stops changing.
  We determined when the distances stopped changing by applying the statistical
  test, \texttt{pymbar.timeseries.detectEquilibration}, to the time series. 
  Simulations of 400 ns give at least 50 ns of equilibrated simulation trajectory.
  %BJC2: Do I cite the github repository for pymbar?
  %MRS4, yes, though also http://pubs.acs.org.colorado.idm.oclc.org/doi/abs/10.1021/acs.jctc.5b00784
  To calculate the equilibrated pore spacing, we measured the distance between
  pore centers. Pore centers are located by averaging the coordinates of sodium
  ions in their respective pores. Pore spacing statistics were generated 
  using the bootstrapping technique (See Supplemental Information).

@@

  % BJC: To be replaced by Joe's description
  {\color{red}This section replaced by Yelk/Glaser text}. 
  Simulated X-ray diffraction patterns are generated based on atomic
  coordinates in order to make a direct experimental comparison. All atomic
  coordinates were simulated as Gaussian spheres of electron density
  corresponding to each atom's atomic number. A three dimensional Fourier
  transform (FT) of the array of electron density results in a three dimensional
  structure factor which represents the unit cell in reciprocal space. We matched
  experimental 2D WAXS patterns by adjusting the initial spacing between layers
  and the orientation of the head groups with respect to adjacent layers.

  The colorbars on all diffraction patterns are normalized relative to
  R-alkanes. We believe that the alkane-alkane density, averaged over all angles,
  is the feature most likely to be replicated between experiment and simulation.
  Other features are dependent on system ordering which is likely to have some
  dependence on initial configuration. We calculated the average intensity within
  R-alkanes of the experimental pattern, $I_{avg}$.  We exclude intensities
  within $\pm$ 30\degree~of the $q_r=0$ axis, since the simulated patterns differ
  from experiment in those regions. 
%MRS4: added to connect.
Specifically, in contrast to the experimental WAXS pattern,
  R-$\pi$ appearing in simulated diffraction patterns intersects with R-alkanes
  (See Fig. \ref{fig:XRDsim}). We multiplied $I_{avg}$ by a scaling factor of
  2.5. Intensities that appear in the experimental pattern $\geq$ 2.5$I_{avg}$
  receive colorbar values of 1. We apply the same scaling method to the simulated
  patterns. We carefully chose a scaling factor of 2.5 in order to (1) visibly
  display all features in the experimental pattern and (2) to allow us to compare
  the relative intensities of features between simulated and experimental
  diffraction patterns.
